---
title: 'Exercise 5_1_1: R Markdown'
author: 'Johannes Breuer, Stefan Jünger, & Veronika Batzdorfer'
date: 'Introduction to R for Data Analysis'
---

```{r child = "./content/config/exercises_setup.Rmd"}
```

The structure of this final exercise for the course is a bit different. We want you to create and work with `R Markdown` documents (creating `HTML` output) to go through some of the things we covered and did in the previous sessions. There are no coding tasks in this document.

We have created an `R Markdown` document (which we have also already knitted to create an `HTML` output file) that demonstrates some of the things you can do/create with `R Markdown` and repeats a few of the topics and steps we went through in the sessions before this one.

This document uses *Gapminder* data and you can find it in the `exercises` folder: It is called `explore_gapminder.Rmd`.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
The first thing we want you to do is to open the `explore_gapminder.Rmd`file in *RStudio* and explore it a bit to see what it contains. You can also open `explore_gapminder.html` (in your browser) to see the `.Rmd` and the resulting output document side-by-side.
```

```{block, opts.label = "clues"}
You can open the `.Rmd` file via the `File` tab or the menu (`File` -> `Open File`) in *RStudio*.
```

You might notice that there are quite a few things specified in the `YAML` header. Let's briefly go through them:

toc: true -> The document will contain a table of contents (ToC)

toc_depth: 3 -> The ToC will contain header levels 1 to 3

number_sections: true -> The sections divided by headers will be numbered

toc_float: true -> The ToC is floating, meaning that it moves when you scroll

code_folding: hide -> By default, the code chunks are hidden, but you display them by clicking the `Code` buttons in the `HTML` document

theme: flatly -> The [*Bootswatch*](https://bootswatch.com/) them [*flatly*](https://bootswatch.com/flatly/) is used to style the document

highlight: tango -> The document uses the [*Pandoc* code highlighting style](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/) *tango*

code_download: true -> The document includes a button allowing you to download the full code

df_print: paged -> When data frames are printed in the document they are printed in paged tables

**NB**: To knit the document you need to have the packages it uses installed. These are the following ones:
`rmarkdown`, `knitr`, `tidyverse`, `visdat`, `janitor`, `pander`, `patchwork`, `correlation`, `GGally`, `broom`, `sjPlot`, `scales`.

An easy option for checking whether you have these packages installed, doing so if that is not the case in one go, and loading them is the `packages()` function from the `easypackages` package. To use it for this purpose, you can run the following code:

```{r easyp, eval = F}
if (!require(easypackages)) install.packages("easypackages")
library(easypackages)

packages("rmarkdown", "knitr", "tidyverse", "visdat", "janitor", "pander", "patchwork", "correlation", "GGally", "broom", "sjPlot", "scales", prompt = F)
```

Feel free to play around a bit with the `explore_gapminder.Rmd` and its output (you can also change parts of the `YAML` header to see how that influences the output).

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
The big task we have for you for this exercise is to create a similar `R Markdown` document for a subset of the *German General Social Survey - ALLBUS 2021* data `gesis_2021_subset`. Using the `explore_gapminder.Rmd` as a starting point and guidance, we want you to do the following in this document:

1. Load and wrangle the data the same way as before to create the `allbus` subset regarding xenophobic attitudes in the exploratory data anlaysis (e.g., select variables like xenophobic attitudes, contact, demographic variables, handle missing values, recode values, and calculate aggregate measures), but with one difference: Also include the variable `ls01` (life satisfaction), rename it to `life_satf`.

2. In addition to that, define a function called `inverter` as follows and use it to create a new variable called `distrust_people` based on `trust`:
`inverter2 <- function (var) {max(var, na.rm = TRUE) - var + 1}`

*Note*: You can do both of these first things in a code chunk that is not displayed in the output document.

3. Get an overview of the missing data using the `vis_miss()` function from the `visdat` package.

4. Look at the relative frequencies for the variables `sex`, `agec`, `educ`, and `party_vote` using a function from the `janitor` package.

5. Create bar plots with `ggplot2` to visualize the relative frequencies (percentages) for the variables `educ` and `party_vote`.

6. Using the `pander` package, include a table with the output of the `base R` function `summary()` for the variables on xenophobic attitudes.

7. Create `ggplot2` bar plots to visualize the distribution of the variables `life_satf` and `left_right`.

8. Create a `ggplot2` boxplot to show differences in xenophobic attitudes supporters of different parties (party_vote); also showing (jittered) individual data points.

9. Calculate correlations between xenophobia, `left_right`, `life_satf` and `xenophobia` using the `correlation` package and display them in a table using the `kable()` from the `knitr` package. 

10. Create a plot with the `GGally` package to visualize these correlations.

11. Calculate a logistic regression model with `trust` as the dependent variable and `contact`, `life_satf`, and `xenophobia` as predictors (also include an intercept).

12. Turn the output of this model into a table with the `broom` package and display it with `knitr::kable()`.

13. Create one regression plot with the coefficients and another one with the predictions using the `sjPlot` package.

This is a lot, but you can find template code (`explore_gapminder.Rmd`) for most of this in the the `solutions` folder (and the rest in the slides for the previous sessions). 
To get you started we have created an almost empty template `.Rmd` called `explore_gesis_allbus_template.Rmd` in the `exercises` folder.
```

```{block, opts.label = "clues"}
If you're stuck, you can find the solutions in the `explore_gesis_allbus.Rmd` in the `solutions` folder.
```
