---
title: 'Exercise 2_2_2: Missing values'
author: 'Johannes Breuer, Stefan Jünger, & Veronika Batzdorfer'
date: 'Introduction to R for Data Analysis'
---

```{r child = "content/config/exercises_setup.Rmd"}
```

In the following exercises, we will explore some of the options for dealing with missing values. As before, we will focus on functions from the `tidyverse` (more specifically, from `dplyr` plus one `tidyr`).   

If you have not yet done so in your current `R` session, please load the same packages that we have used in the previous set of exercises (`sjlabelled`, `tidyverse`, `haven`; ideally also in that order).

```{r libraries, include=FALSE}
library(sjlabelled)
library(tidyverse)
library(haven)
```

We should also import the data again.

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Import the *ALLBUS* 2021 `.sav` and assign it to an object named `allbus_2021_missings`. However, unlike in the previous set of exercises, for our missing value operations, we do not want to user-defined missing values from the *SPSS* file to be converted to `NA` when the data are imported.
Same as before, we want to remove all labels.
```

```{block, opts.label = "clues"}
To override the default behavior of converting user-defined missing values into `NA`, we need to provide an appropriate value for the `user_na` argument of the import function we need. To check what the appropriate value is, you can look at the function help file (`?read_sav()`).
```

```{r first-exercise, solution = TRUE}
allbus_2021_missings <- read_sav("./data/allbus_2021/ZA5280_v1-0-0.sav",
                        user_na = FALSE) %>% 
  remove_all_labels()
```

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
As a first exercise, let's turn all negative values for the variable measuring satisfaction with democracy in Germany into `NA`s using a function from the `dplyr` package. 
```

```{block, opts.label = "clues"}
To to this, we need to combine `mutate()` with the `dplyr` function for recoding specific values as `NA`. The name of the variable is `ps03` and the values we need to turn into `NA`s are: -42, -11, -9, -8.
```

```{r second-exercise, solution = TRUE}
allbus_2021_missings <- allbus_2021_missings %>% 
  mutate(ps03 = na_if(ps03, -42)) %>% 
  mutate(ps03 = na_if(ps03, -11)) %>% 
  mutate(ps03 = na_if(ps03, -9)) %>%
  mutate(ps03 = na_if(ps03, -8))
```

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
After recoding a set of values as `NA` for one variable, let's now do the same for a whole dataframe. For this purpose, first create a new dataframe named `contact` that only contains variables assessing in which contexts respondents have contact with immigrants. Then convert the following values to `NA` for all variables in this dataframe: -42, -11, -10, -9.
```

```{block, opts.label = "clues"}
The names of the variables we are looking for are: `mc01`, `mc02`, `mc03`, and `mc04`. For converting values to `NA` for all columns/variables in a dataframe, we do not need the `mutate()` function.
```

```{r third-exercise, solution = TRUE}
contact <- allbus_2021_missings %>% 
  select(mc01:mc04)

contact <- contact %>% 
  na_if(-42) %>% 
  na_if(-11) %>% 
  na_if(-10) %>% 
  na_if(-9)
```

```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
As `na_if()` only takes only takes single values as its second argument (i.e., the value to replace with `NA`), let's use a function from the `sjlabelled` function to achieve the same thing as in the previous task with fewer lines of code.
```

```{block, opts.label = "clues"}
The function we are looking for can also be included in a pipe chain and takes a vector of values to be recoded as `NA` as its second (required) argument. 
```

```{r fourth-exercise, solution = TRUE}
library(sjlabelled)

contact <- contact %>% 
  set_na(na = c(-42, -11, -10, -9))
```

```{block, box.title = "5", box.body = list(fill = "white"), box.icon = "fa-star"}
As we final exercise, let's build on what we did in task 2 and use a function from the `tidyr` package to answer the following question: How many of the respondents do not have a missing value for the item assessing satisfaction with democracy in Germany? This time, do not assign the result of your code to a new object.
```

```{block, opts.label = "clues"}
To count the number of cases, you can use the base `R` function `nrow()` at the end of your pipe.
```

```{r fifth-exercise, solution = TRUE}
allbus_2021_missings %>% 
  drop_na(ps03) %>% 
  nrow()
```
