---
title: "Exploring data from the GESIS German General Social Survey - ALLBUS 2021"
author: "R User"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: flatly
    highlight: tango
    code_download: true
    df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

options(scipen = 15)
```

#Data Wrangling

```{r load-wrangle-data, echo = TRUE}
library(tidyverse)
library(naniar)
library(sjlabelled)
library(haven)

inverter <- function (var) {
  max(var, na.rm = TRUE) - var + 1
}

allbus_2021 <- read_sav("../data/allbus_2021/ZA5280_v1-0-0.sav") %>% 
  remove_all_labels()

allbus_2021_subset  <- allbus_2021  %>% 
  select(respid,
         sex,
         agec,
         educ,
         left_right = pa01,
         party_vote = pv01,
         sat_dem = ps03,
         xeno1 = ma01b,
         xeno2 = ma02,
         xeno3 = ma03,
         xeno4 = ma04,
         contact1 = mc01,
         contact2 = mc02,
         contact3 = mc03,
         contact4 = mc04,
         trust_gov = pt12,
         life_satf = ls01
  ) %>% 
  replace_with_na_all(condition = ~.x < 0)  %>% 
  mutate(sex = recode_factor(sex,
                             `1`= "Male",
                             `2` = "Female",
                             `3` = "Non-binary"),
         sat_dem = recode(sat_dem,
                          `1` = 6,
                          `2` = 5,
                          `3` = 4,
                          `4` = 3,
                          `5` = 3,
                          `6`= 1),
         agec = recode_factor(agec,
                              `1`= "<= 25 years",
                              `2`= "26 to 30 years",
                              `3` = "31 to 35 years",
                              `4` = "36 to 40 years",
                              `5` = "41 to 45 years",
                              `6` = "46 to 50 years",
                              .ordered = TRUE),
         party_vote = recode_factor(party_vote,
                                    `1`= "CDU-CSU",
                                    `2`= "SPD",
                                    `3` = "FDP",
                                    `4` = "Gruene",
                                    `6` = "Linke",
                                    `42` = "AfD",
                                    `90` = "Other party",
                                    `91` = "Would not vote"),
         distrust_gov = inverter(trust_gov)) %>% 
  mutate(
    xenophobia = rowMeans(across(
      xeno1:xeno4))) %>% 
  mutate(across(contact1:contact4,
                ~recode(
                  .x,
                  `2` = 0))) %>% 
  mutate(contact = rowSums(across(
    contact1:contact4)))

```

# The data

The data we use comes from the [German General Social Survey - ALLBUS 2021 ](https://search.gesis.org/research_data/ZA5280).We have selected `r ncol(allbus_2021_subset)` variables of interest for our analysis. The dataset includes data for *N* = `r nrow(allbus_2021_subset)` respondents.   


## Missing data

As not all of the respondents answered every question in the survey, we have missing data. Before we further explore the data, we want to know whether there are identifiable patterns of missingness in them.

```{r vis-miss}
library(visdat)

vis_miss(allbus_2021_subset)
```

# Exploring the *GESIS German General Social Survey - ALLBUS 2021* dataset

## Demographics

We first want to look at gender, age (categories), education, and party vote in our sample.

```{r demo-tables}
library(janitor)

allbus_2021_subset %>% 
  tabyl(sex) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting() %>% 
  knitr::kable()

allbus_2021_subset %>% 
  tabyl(agec) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting() %>% 
  knitr::kable()

allbus_2021_subset %>% 
  tabyl(educ) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting() %>% 
  knitr::kable()

allbus_2021_subset %>% 
  tabyl(party_vote) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting() %>% 
  knitr::kable()

allbus_2021_subset %>% 
  ggplot(aes(x = educ, 
             fill = educ)) + 
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels=scales::percent) +
  ylab("Relative Frequencies")

allbus_2021_subset %>% 
  filter(!is.na(party_vote)) %>% 
  ggplot(aes(x = party_vote, 
             fill = party_vote)) + 
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels=scales::percent) +
  ylab("Relative Frequencies")

```

## Summary statistics

Let's look at the summary statistic for the items on xenophobia.

```{r summarystats}
library(pander)

allbus_2021_subset %>% 
  select(starts_with("xeno")) %>% 
  summary() %>% 
  pander()
```

## Distribution of variables

We want to explore how some of the data for the question on life satisfaction and political orientation (left-right) are distributed in our sample.

```{r density-satf}
allbus_2021_subset %>% 
  ggplot(aes(x = life_satf)) +
  geom_bar() +
  scale_y_continuous(expand=expansion(mult=c(0,0.1)))
```

```{r density-lr}
allbus_2021_subset %>% 
  ggplot(aes(x = left_right)) +
  geom_bar() +
  scale_y_continuous(expand=expansion(mult=c(0,0.1)))
```

## Party affiliation and xenophobic attitudes

How do xenophobic attitudes differ between supporters of different parties? 

```{r party-xenoph}
allbus_2021_subset %>% 
  ggplot(aes(x = party_vote, y = xenophobia)) +
  geom_boxplot() +
  geom_jitter(position = position_jitter(width = 0.1, height = 0),
              alpha = 0.25)
```

# Analyses

## Correlations

We want to explore the correlations between political leaning, xenophobia, and life satisfaction.

```{r correlations}
library(correlation)
library(GGally)

allbus_2021_subset %>% 
  select(left_right,
         xeno1:xeno4,
        life_satf) %>% 
  correlation %>% 
  knitr::kable()

allbus_2021_subset %>% 
  select(left_right,
         xeno1:xeno4,
         life_satf) %>% 
  ggcorr(label = TRUE,
         label_round = 2)
  
```

## Regression analysis

We use a logistic regression model to explore how trust in people is predicted by life satisfaction, mean xenophobic attitudes and the contact with foreigners.

```{r regression}
allbus_2021_subset <- allbus_2021_subset %>% 
  mutate(distrust_gov_dic = ifelse(distrust_gov > 4, 0, 1))

library(broom)

model <- glm(distrust_gov_dic ~ 1 + life_satf + contact + xenophobia,
             family = binomial(link = "logit"),
             data = allbus_2021_subset)

tidy_model <- model %>% 
  tidy()

tidy_model %>% 
  knitr::kable()

glance(model)
```

The unstandardized regression coefficients *b* can be selected from the `tidy_model`.

### Regression plots

```{r regression-plots}
library(sjPlot)

plot_model(model)

plot_model(model,
           type = "pred")
```
