---
title: "Introduction to R for Data Analysis"
subtitle: "Outlook"
author: "Johannes Breuer, Stefan Jünger, Veronika Batzdorfer"
date: "2021-08-19"
presenter: Stefan
editor_options: 
  chunk_output_type: console
---

layout: true 

```{r child = "./content/config/sessions_setup.Rmd"}
```

---

## Recap: Course schedule - Day 1

```{r schedule-1, echo = F}
course_content_1
```

---

## Recap: Course schedule - Day 2

```{r schedule-2, echo = F}
course_content_2
```

---

## Recap: Course schedule - Day 3

```{r schedule-3, echo = F}
course_content_3
```

---

## Recap: Course schedule - Day 4

```{r schedule-4, echo = F}
course_content_4
```

---

## Recap: Course schedule - Day 5

```{r schedule-5, echo = F}
course_content_5
```

---

## Our jou`R`ney this week (hopefully)

```{r mindblowing, echo = FALSE, out.width = "85%"}
woRkshoptools::include_picture("r_pkgs_mindblowing.png")
```
.footnote[[Source](https://res.cloudinary.com/syknapptic/image/upload/v1521320144/tidyverse_meme_oceake.png)]

---

## Where to go from here?

Hopefully, after this week, you feel prepared to do your next steps in `R`.

Some recommendations for continuing your jou`R`ney:

- Keep up working with `R`!
- If time permits, do stuff you usually do in `SPSS` or `Stata` in `R`, even when it's harder
- Try to do at least one research task solely in `R` (one analysis, a whole paper, a report, etc.)
- Look for tutorials and guides online
  - trust us, there's way more (good & free) online material for `R` than there is, e.g., for `SPSS` or `Stata`

---

## Resources: Introductory books

[R for Data Science](https://r4ds.had.co.nz/) by Hadley Wickham

[R Cookbook: Proven recipes for data analysis, statistics, and graphics](https://rc2e.com/) by JD Long & Paul Teetor

[Hands-On Programming with R](https://rstudio-education.github.io/hopr/) by Garrett Grolemund

[R Programming for Data Science](https://bookdown.org/rdpeng/rprogdatascience/) by Roger D. Peng

[Quantitative Social Science Data with R](https://uk.sagepub.com/en-gb/eur/quantitative-social-science-data-with-r/book257236) by Brian J. Fogarty

[Introduction to R for Social Scientists - A Tidy Programming Approach](https://www.routledge.com/Introduction-to-R-for-Social-Scientists-A-Tidy-Programming-Approach/Kennedy-Waggoner/p/book/9780367460723) by Ryan Kennedy & Philip D. Waggoner

[Quantitative Social Science: An Introduction in tidyverse](https://press.princeton.edu/books/hardcover/9780691222271/quantitative-social-science) by Kosuke Imai & Nora Webb Williams

---

## Resources: Online courses & tutorials

- Overview of resources
  - [*learnR4free*](https://www.learnr4free.com/) by Mine Dogucu
  - the [*Big Book of R*](https://www.bigbookofr.com/) by Oscar Baruffa

- [*swirl* - Learn `R` in `R`](https://swirlstats.com/)

- One of my personal favorites: Learning `R` (and statistics) with a cute story and beautiful illustrations: [Teacups, Giraffes, & Statistics by Hasse Wallum & Desirée de Leon](https://tinystats.github.io/teacups-giraffes-and-statistics/)

---

## Working with other data types

Johannes, Stefan, and Veronika use different data types in their daily work:
- digital trace data (Johannes/Veronika) <sup>*</sup>
- georeferenced/geospatial data (Stefan)<sup>**</sup>

**Remember that `R` is data-agnostic! It can serve as a fancy data science tool for extracting social media data but also as a full-blown Geographic Information System (GIS)**


.small[
<sup>*</sup> see, e.g., https://github.com/jobreu/twitter-linking-workshop-2022 or https://github.com/jobreu/youtube-workshop-gesis-2022

<sup>**</sup> see, e.g., https://github.com/StefanJuenger/gesis-workshop-geospatial-techniques-R or https://github.com/StefanJuenger/esra-workshop-first-steps-R-GIS
]

---

## Working with text data in `R`

As with almost everything else, there are many great resources for working with text data in `R`. Two good options (and starting points) are:
- the [`tidytext` package](https://juliasilge.github.io/tidytext/) and the "accompanying" book [*Text Mining with R: A Tidy Approach*](https://www.tidytextmining.com/) by Julia Silge & David Robinson
- the [`quanteda` package](https://quanteda.io/) and its accompanying [tutorials](https://tutorials.quanteda.io/)

- a nice and free self-paced online course is [*Text mining in R for the social sciences and digital humanities*](https://tm4ss.github.io/docs/index.html) by Andreas Niekler and Gregor Wiedemann

---

## What Are Geospatial Data?

.pull-left[
Data with a direct spatial reference $\rightarrow$ **geo-coordinates**
- Information about geometries
- Optional: Content in relation to the geometries

Can be projected jointly in one single space
- Allows data linking and extraction of substantial information
]

.pull-right[
```{r pic-geometries-1, echo = FALSE, out.width = "85%"}
woRkshoptools::include_picture("fig_geometries.png")
```

.tinyisher[Sources: OpenStreetMap / GEOFABRIK (2018), City of Cologne (2014), and the Statistical Offices of the Federation and the Länder (2016) / Jünger, 2019]
]

---

## Mapping is so easy nowadays

.pull-left[
```{r mapsf, eval = FALSE}
library(mapsf)

mtq <- mf_get_mtq()

mf_map(x = mtq)
mf_map(x = mtq, var = "POP", type = "prop")
mf_layout(
  title = "Population in Martinique", 
  credits = "T. Giraud; Sources: INSEE & IGN, 2018"
)
```
]

.pull-right[
```{r mapsf-print, echo = FALSE}
library(mapsf)

mtq <- mf_get_mtq()

mf_map(x = mtq)
mf_map(x = mtq, var = "POP", type = "prop")
mf_layout(
  title = "Population in Martinique", 
  credits = "T. Giraud; Sources: INSEE & IGN, 2018"
)
```
]

Example from: https://riatelab.github.io/mapsf/


---

## 'Web development' using `R`

These days, a lot of `R` packages provide tools originally developed for the web. For example:
- [bookdown](https://pkgs.rstudio.com/bookdown/) enables you to publish books written in `R Markdown` online
- [pkgdown](https://pkgdown.r-lib.org/) does the same for your own `R` packages
- [blogdown](https://pkgs.rstudio.com/blogdown/l) is more general and helps you with creating websites (examples to follow)

---

## Shiny apps

> Shiny is an R package that makes it easy to build interactive web apps straight from R. You can host standalone apps on a webpage or embed them in R Markdown documents or build dashboards. You can also extend your Shiny apps with CSS themes, htmlwidgets, and JavaScript actions.

https://shiny.rstudio.com/

---

class: middle

## Example 1: Movie Explorer

.center[https://shiny.rstudio.com/gallery/movie-explorer.html]

---

class: middle

## Example 2: CRAN explorer

.center[https://gallery.shinyapps.io/cran-explorer/]

---

## Creating your own homepage with `R`

.pull-left[
```{r home-johannes, echo = FALSE}
woRkshoptools::include_picture("homepage_johannes.png")
```
.center[.small[https://www.johannesbreuer.com/]]
]

.pull-right[
```{r home-stefan, echo = FALSE}
woRkshoptools::include_picture("homepage_stefan.png")
```
.center[.small[https://stefanjuenger.github.io/]]
]

.center[Powered by  [`blogdown`](https://cran.r-project.org/web/packages/blogdown/index.html) &[ Hugo Academic](https://academic-demo.netlify.app/)]

---

## Writing your own `R` packages

.pull-left[
At a certain point (not now!), you may want to consider writing your own `R` package
- useful for creating reproducible code
- great for distributing your work to others
- for example, we created an [`R` package](https://stefanjuenger.github.io/woRkshoptools/) to facilitate working on our workshop materials
]

.pull-right[
```{r r-packages, echo = FALSE}
woRkshoptools::include_picture("r_packages.jpg")
```
[Read the book here!](https://r-pkgs.org/)
]

---

class: middle

## It's straightforward in `RStudio`

```{r new-package, echo = FALSE, out.width = "75%"}
woRkshoptools::include_picture("new_package.png")
```

---

## Acknowledgements `r ji("heart")`

All slides were created with the `R` package [`xaringan`](https://github.com/yihui/xaringan) which builds on [`remark.js`](https://remarkjs.com), [`knitr`](http://yihui.name/knitr), and [`R Markdown`](https://rmarkdown.rstudio.com). The exercises were created with the [`unilur` package](https://github.com/koncina/unilur).  

Please make sure to properly cite all data that you use for your research (archives usually provide suggested citations). Also make sure to cite the free and open-source software (FOSS) that you use, such as `R` and the packages for it.

Veronika, Stefan, and Johannes want to thank Theresa for her excellent support before and during the course!

Finally, all of us want to thank the *GESIS Training* team for taking good care of the organization of this course (and the whole Summer School) and all of you for participating!
